---
title: "Standard Errors in Regression Coefficients"
output: github_document
---

> Standard Error is the standard deviation of a sampling distribution

Uncertainty in the estimated regression coefficients is given by standard errors (i.e. how much, on average, could estimated coefficients vary from the true population coefficients). 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)

set.seed(7)

employees <- data.table(IsFaculty = rep(x = c(0,1), each = 100),
                        ServiceYears = c(rnorm(n = 100, mean = 5, sd = 2), 
                                         rnorm(n = 100, mean = 7, sd = 2)))

lm.employees <- lm(formula = ServiceYears ~ IsFaculty, data = employees)

lm.employees.tidy <- lm.employees %>% broom::tidy() %>% as.data.table()

lm.employees.tidy
```

In this example regressing ServiceYears on whether the employee is Faculty (1) or staff (0), repeated sampling of the cofficient estimates would vary, on average, +/- .27 service years around the true population coefficient. The interval from -.27 + 1.98 to .27 + 1.98 (+/- one SE) would contain the true population coeficient of IsFaculty roughly 68% of the time; ~95% for +/- two SE; ~99.7% for three SE etc.

Below, the regression line is shown with +/- two standard errors (blue). Since this area does not contain
the intercept (orange) we can conclude, with roughly 95% confidence, that the coefficient estimate
is significantly non-zero and somewhere between `r (-.27*2) + 1.98` and `r (.27*2) + 1.98` in the population.

```{r}
#Adjust fitted values to reflect +/- two standard errors of estimated coefficient IsFaculty and assign to employees
employees[, ":="(
  TwoSELower = lm.employees.tidy[term == "(Intercept)", estimate] +
    ((lm.employees.tidy[term == "IsFaculty", estimate] - 
      (2 * lm.employees.tidy[term == "IsFaculty", std.error])) * IsFaculty),
  TwoSEUpper = lm.employees.tidy[term == "(Intercept)", estimate] + 
    ((lm.employees.tidy[term == "IsFaculty", estimate] + 
      (2 * lm.employees.tidy[term == "IsFaculty", std.error])) * IsFaculty)
  )
]

ggplot(data = employees, aes(x = IsFaculty, y = ServiceYears)) +
  geom_point(alpha = .25) +
  geom_line(aes(y = lm.employees$fitted.values)) +
  geom_ribbon(aes(ymin = TwoSELower, ymax = TwoSEUpper),
              alpha = .5, 
              fill = "slategray2") +
  geom_line(aes(y = lm.employees.tidy[term == "(Intercept)", estimate]), 
            linetype = 2, 
            color = "orange")
```