---
title: "Understanding Regression Coefficients with One Predictor"
output: github_document
---

### Binary Predictors

> The regression coefficient is the difference between the averages of the two groups.

Suppose a regression of Salary on whether the employee is class Faculty (1) or Staff (0):

![](http://www.sciweavers.org/tex2img.php?eq=%20Salary%20%3D%20%20%5Cbeta_0%20%2B%20%5Cbeta_1%20IsFaculty%20%2B%20e&bc=White&fc=Black&im=png&fs=12&ff=modern&edit=0)

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(package = data.table)
library(package = tidyverse)

set.seed(7)

salaries <- data.table(IsFaculty = rep(x = c(0, 1), each = 50),
                       Salary = c(rnorm(n = 50, mean = 50000, sd = 10000),
                                  rnorm(n = 50, mean = 60000, sd = 10000)))

lm.salaries <- lm(formula = Salary ~ IsFaculty, data = salaries)

lm.salaries$coefficients %>% round(0)
```

The intercept represents the predicted value when all predictors are 0 (i.e. Staff). The regression coefficient tells us that, in our sample, Faculty make $8k more in Salary on average than Staff.  

An unbiased linear model intersects the average of each sub-population:

![](http://www.sciweavers.org/tex2img.php?eq=%20%5Cwidehat%7BSalary%7D%20%3D%20%2052%2C387%20%2B%208%2C000%20%5Ctimes%20IsFaculty&bc=White&fc=Black&im=png&fs=12&ff=modern&edit=0)

```{r}
ggplot(data = salaries, aes(x = IsFaculty, y = Salary)) + 
  geom_point(alpha = .25) +
  geom_line(aes(x = IsFaculty, y = lm.salaries$fitted.values))
```

### Continuous Predictors

> The regression coefficient is the difference between the average of sub-populations 1 unit apart.

or

> The regression coefficient is the slope of the prediction line at any level.

Suppose a regression of Salary on Age in continuous years:

![](http://www.sciweavers.org/tex2img.php?eq=Salary%20%3D%20%20%5Cbeta_0%20%2B%20%5Cbeta_1%20Age%20%2B%20e&bc=White&fc=Black&im=png&fs=12&ff=modern&edit=0)

```{r}
set.seed(7)

salaries <- data.table(Age = sample(x = 18:69, 
                                    size = 500, 
                                    replace = TRUE)) %>%
  mutate(Salary = Age * 1000 + rnorm(n = 500, mean = 10000, sd = 10000))

lm.salaries <- lm(formula = Salary ~ Age, data = salaries)

lm.salaries$coefficients %>% round(0)
```

The intercept represents the predicted value when all predictors are 0 (i.e. Salary at birth--not very meaningful). The regression coefficient tells us that, in our sample, employees make $1,019 more Salary on average than employees one year their junior; or you could say that an employee would make \$10,190 more Salary on
average if they were a decade older.  

![](http://www.sciweavers.org/tex2img.php?eq=%20%5Cwidehat%7BSalary%7D%20%3D%20%209%2C145%20%2B%201%2C019%20%5Ctimes%20Age&bc=White&fc=Black&im=png&fs=12&ff=modern&edit=0)
```{r}
ggplot(data = salaries, aes(x = Age, y = Salary)) +
  geom_point(alpha = .25) +
  geom_line(aes(x = Age, y = lm.salaries$fitted.values))
```